Alleviating the Topology Mismatch Problem in Unstructured and Structured P2P Overlay Networks: A Survey

### 2014-01-23 background.tex
2012.06.26 (Mema) Big introduction section split to a crispier 3-page intro and a background section. The latter also shortened and updated.

\section{Introduction}
\label{section:intro}

Distributed architectures have witnessed a significant increase of interest
during the last decade, as the number of interconnected computers increased, the
available digital information kept proliferating, at an exponetial rate each
year, and problems to be addressed by IT became more and more demanding. Why is
there a correlation, and it is not just a coincidence? Distributed computation
suggests a simple paradigm\footnote{Of course, as always, this is the case only
after someone thought about it first!}. Get a set of autonomous computing
entities and put them to cooperate with each other in order to fullfil a well
defined task. In the World of the Internet and the World Wide Web this also
offers a thrilling promise. To harness the immence power scattered across the
Internet and get amply rewarded by exploiting the data, storage and
computational resources out there.

In order to achieve that, computers collaborate by exchanging messages via an
application-level network where they can address each other in the specific
context of the task at hand. This communication structure is refered to as an
\emph{overlay network} which is an abstract, logical inter-connection formed on
top of another network\footnote{The Internet itself is an overalay on top of the
telephone network.}. It consists of a set of nodes, the computing elements, that
are being connected by virtual (logical) point-to-point links, perhaps
corresponding to multiple physical links. Figure~\ref{figure:overlay}
illustrates a simple four-node overlay contructed upon a wide area network.

\begin{figure}
\centering
  \includegraphics[scale=0.5]{img/p2p.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

The abstraction of the overlay network has been proposed as a way to implement
efficient, fully distributed, application-layer services. For example, backing
the routing of messages to destinations that are not known in advance or
building one-to-many addressing schemes
% TODO find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
\cite{} that would provide higher quality-of-service (QoS) guarantees. \emph{IP
multicast}, as well as other proposals such as \emph{IntServ} and
\emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance
(yet?), largely due to the fact they require support from the physical
infrastructure. On the other hand, an overlay network can be deployed on
end-systems running the overlay protocol software, without cooperation from the
ISPs.
%Academic research includes
%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peer-to-Peer (P2P) Architectures in a Nutshell}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Overlay networks on top of the best-effort IP layer, formed a widely ranged
family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for
short. P2P systems consist of a number of networked computers called
\emph{peers}, where each, such, \emph{peer} could act both as a resource
provider and a resource consumer. This changed the traditional
\emph{client-server} model dominating the Internet and lead to the introduction
of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends
the notions of server and client to denote the twofold role of the P2P network
participants. Such serverless systems, proved to be able to achieve outstanding
aggregate resource capacities as more and more participants join the system
without requiring additional expenditure for infrastructure\footnote{
  Unfortunatelly, there has been observed mitigation of this self-scaling
  property by the undesirable behaviour of participating users, usually answered
  in popular file-sharing systems; from \emph{free-riding}

\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005}
to the distribution of illegal content and/or other \emph{socio-technical}
issues \cite{hughes_socp2p_2008}.
}.
P2p network systems gained extra attention whithin the distributed systems
super-set, circa 2000, due to the fact that they succesfully supported popular
applications that offered file sharing among vast numbers of participating
users. But their development didn't stay still. On the contrary, during the last
decade, research on P2P systems was constant and fruitfull generating a wide
range of popular protocols, networks and applications. As P2P systems evolved
three main architectures have been emerged, namely
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{centralized},
  \item \emph{decentralized unstructured}, and
  \item \emph{decentralized structured}.
\end{inparaenum}

\emph{Centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised})
were the first to recognize that requests for resourses (e.g. CPU cycles
\cite{seti} or popular content
% TODO find citation for Napster
\cite{}) need not be sent to a dedicated server but instead could be handled by
the many hosts that already posses them. This schema, though, is not fully
distributed since it dictates a centralised searching facility, a characteristic
that can be proven its ``Achilles' heel'' in terms of being a scalability
bottleneck, a central point of failure and vulnerable to malicious acts (e.g.
\emph{Denial-of-Service (DoS) attacks}).

The centralized approach has been eventually replaced by architectures that
distributed both search and download capabilities (see
Figures~\ref{figure:p2p-archs:unstructured-full} and
\ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized}
architectures, file placement is random, which means there is no correlation
with the network topology whatsoever \cite{yang_improvep2psearch_2002}. For this
reason they are also refered to as \emph{unstructured}. The most important
properties of such systems are that they support inherent heterogeneity of
peers, are highly resilient to peer failures, and incur low maintenance overhead
at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In
the literature, they are also known as \emph{broadcast-based} systems, because
they use \emph{message flooding} among peers or super-peers to propagate
queries.
%\footnote{The life of each query message is constrained by an integer, refered
%to as \emph{time-to-live (TTL)}. When a querry is received by a (super-) peer,
%this value is decreased by one unit. Then, the receiver, back-propagates its
%results and if the TTL has not reached zero the node forwards the query message
%to all its neighbours except from the one it received it from.}

Recently, \emph{decentralized structured} schemes have been proposed (see
Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising
infrastructure for large-scale P2P applications
\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
through a deterministic mechanism. Even though unstructured networks became
highly popular among file sharing applications, they do have a certain
limitation in locating rare objects. The DHT approach of structured networks on
the other hand, provides a guaranteed bound on the number of overlay routing
hops that have to be taken in order for an object to be located, even in the
case where only a single copy exists in the system.  This means
$O \left ( log n \right )$ hops, compared to unstructured networks that require
$O \left ( n \right )$ to reliably locate a specific object\footnote{
  Unfortunatelly this efficiency does not come without cost. Structured overlays
  realise what is called object identifier location as opposed to content-based
  retrieval of unstructured ones. This means that the later can support more
  versatile object location mechanisms like keyword and partial matching
  searches.}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Applications and Implications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The most successful incarnation of the centrilised approach is without a doubt
the popular file-sharing application \emph{Napster}\footnote{Pioneering, though,
was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare
processing power of participants to search the universe for the existence of
extra terrestrial intelligence!}
% TODO find citation for Napster again
\cite{}.
Napster maintained a \emph{central index server} based on file lists provided by
participating peers. The central index server was queried by the users and it
returned pointers to the actual content. Thus, by centralizing search while
distributing downloads, Napster achieved a highly functional design that was
widely acknowledged, at the time, as ``the fastest growing Internet application
ever'' that toped some 26 million users and 80 million songs
\cite{jmm_naptopusage_2001}.

Napster was ultimatelly brought down after Recording Industry Association of
America (RIAA) filed a lawsuit for copyright infringement of the intellectual
property of artists whose music was distributed by the network\footnote{
  The sueing spree was initiated by heavy metal giants \emph{Metallica} that
  feared Napster was undermining record sales. Opposed opinions instantly
  emerged with best example the british alternative rock band \emph{Radiohead}
  whose album \emph{Kid A} had also been leaked to Napster three months before
  reaching music store shelves \cite{rm_radioheadkida_2000}. Despite music
  critics labeling the album ``experimental'' and ``challenging'' (it had no
  official singles nor music videos for any of its songs), in October 2000 Kid A
  captured the number one spot on the Billboard 200 sales chart in its debut
  week. Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster,
  Radiohead had never hit the top20 in the US before.
}
\cite{wiki_amnaptrial_2001}. This triggered a shift towards decentralised
networks like the classic \emph{Gnutella} and, its ``rival'' at the time,
\emph{FastTrack}. Gnutella, in its first realisation (v.0.4 \cite{gnutellav04}),
was a fully decentralised approach. Unfortunatelly, network resourses tended to
be overwhelmed by message flooding and if queries were constrained to live for
some (small) number of hops, this ultimatelly resulted to a shrinked search
scope, leading to a death spiral that prevented the network from scaling. The
solution came with a slightly more structured network by introducing the notions
of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}). With
\emph{Limewire}, the most widely used software that implemented the protocol,
the network was highly successful and was reported to have reached connecting
more that 3 million users concurrently by the end of 2006
\cite{rsr_gnutellaevol_2006}. Fasttrack and its implementing software KaZaA
\cite{kazaa}, on the other hand, went with the hierarchical approach
(\emph{super-nodes} in their context) from the very begining, dynamically
assigning indexing functions to special peers in order to increase search
performace and scaling potential. FastTrack network had an average of 3 million
concurrently connected nodes in 2003 \cite{lkxr_kazaameasure_2005}. It is
estimated that the total number of users was greater than that of Napster at its
peak. With Limewire discontinued/shutdown in 2010 and its successor,
\emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well
signal the end of the once-mighty Gnutella network. KaZaA continues as a
propriatery music store solution. In the meanwhile, Fastrack design team had
already introduced a VoIP and text messaging application based on the design
principles of KaZaA's hierarchical P2P nature, that was meant to change how
people communicate around the world. This app was no other than Skype
\cite{skype}. Despite its wide usage, little is known for the technical aspects
of its overlay network, mainly due to its proprietary nature. Studies, though,
confirm that it shares strong structural similarities to its file-sharing
predecessor (ordinary hosts and super-nodes in Skype's terminology) but they
significantly differentiate in terms of traffic behaviour
\cite{g_voipskypesec_2005,gdj_skypestudy_2006}. Skype was introduced in 2003
and from then on it witnessed frenzy-rythmed increase to its popularity, with
more that 650 milion registered users in 2011 \cite{skypetotalusers} and
reported to have broken the barrier of 35 million skypers concurrently online on
the the 5th of March 2012 \cite{skypesymusers}. On October 2011 Microsoft Corp.
and Skype Global S.à.r.l. anounnced the completion of Microsoft's acquisition of
Skype for \$$8.5$ billion \cite{skypemicrosoft}.

File sharing was the most popular application area for structured architectures
as well. Overnet
% TODO find citation for Overnet
\cite{},
for example, was created for sharing large files (e.g., movies and CD images).
It implemented the Kademlia tree-like partitioning algorithm. Other
implementations of Kademlia are the Kad Network (eMule/aMule  clients),
BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent,
BitComet, Transmission and BitSpirit all share compatibility with the Mainline
DHT which realises the Kademlia protocol) and the Gnutella DHT (first
implemented by LimeWire as the Mojito DHT). Lately, RetroShare \cite{retroshare}
is the new trend for file-sharing which incorporates an implementation of the
Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of
a darknet as well. But the potential of structured architectures was not
exhausted only to the above. Structured P2P networks became popular in
distributed systems because they could scale smoothly and have fast and reliable
object location mechanisms. Thus, research and development was very active and
productive resulting in a variety of diverse applications such like distributed
search engines \cite{yaci}, distributed data storage systems
\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003}, Web caches, archives,
and publishing systems \cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
messaging applications \cite{threedegrees}, event notification infrastructures
\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, naming services
\cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and
lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.

%%%%%%%% NEW TRENDS %%%%%%%%
%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in
%response to the continued push for more anti-piracy tools, the final Pirate Bay
%verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA. This
%leads to users trying to find new solutions for anonymous, uncensored,
%decentralised file sharing solutions like Freenet, Tribler and RetroShare where
%connections are made only between trusted peers — sometimes called "friends"
%and the respective formed network F2F.

% TODO: READ Dalesa published at eAsia 2009, http://www.e-asia.org/2009/ET_Presentations/Wathsala.pdf

The huge numbers of users, for the various applications we presented above,
reveals the tremendous popularity P2P architectures have enjoyed. This, though,
does not come without any toll. Measurements on some popular systems, such as
FastTrack, Gnutella, DirectConnect and BitTorrent have shown that the P2P
traffic contributes the largest portion of the overall Internet traffic
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}.
Even with later studies, this observation still holds on. Using deep packet
inspection techniques deployed at selected customer sites \emph{Ipoque} and
\emph{CacheLogic} were able to track down application usage \cite{cachelogic,ipoque2007,ipoque2009}.
CacheLogic reports that more than 60\% of the traffic is identified as P2P, with
Ipoque verifying this observation reporting similar results for 2007, while
observing some small decline for 2008/2009 (but still remaining dominant
compared to other applications). Apart from its participation in overall usage
terms, P2P is already a huge contibutor to the overall data exchanged across the
Internet\footnote{
  Mainly because file-sharing P2P is used to exchange big files of HD video or
  high sampled audio content compared to simple Web content distributed via, for
  example, HTTP.
}.
\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes
per month and the potential is to be quintubled by the end of 2012 reaching
almost 8 petabytes per month.

%
% TODO: SOME DISCUSSION from \cite{zhao_ctag_2006}
%
%%Observations made in various studies (e.g. \cite{matei_mapgnutella_2002})
% concerning the distribution of nodes among \emph{Internet Service Providers
%(ISPs)} and \emph{Autonomous Systems (ASs)} have shown that only $2$ to $5$
%percent of Gnutella connections link peers within a single AS while more than
%$40$ percent of all Gnutella peers are located within the top 10 ASs.
%Similarly, measurements in \cite{zeinalipour-yazti_gnudc_2002} used a $244,000$
%IPs test-bed and results have shown that $45$ percent of the nodes belonged to
%only $10$ large ISPs and $58$ percent belong to only $20$. Such results mean
%that most overlay generated traffic crosses AS borders increasing topology
%mismatch cost.

@@

If these peers participate in an overlay network according to one of the setups
of Figure~\ref{figure:overlay-confs} then users will expierience different
performances. In the overlay depicted in Figure~\ref{figure:overlay-1}, the
message will traverse the following sequence of links in the physical layer
(marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2
\rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5
\rightarrow 5 \rightarrow 4$. In the alternative overlay of Figure
~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$. The
total cost of the first sequnce is $45 ms$, while the second yields a mere $15
ms$. Therefore, we can conclude that the second overlay is more congruent with
the underlying physical network than the first one and, thus, is more efficient.

@@

Ideally, overlay networks are expected to achieve an optimal mapping among the
overlay paths and the underlying physical links, and avoid inefficient states.
The problem of constructing an optimal overlay is referred to as the
\emph{topology mismatch problem}, and formally defined as follows:
\begin{definition}
Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
$\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
$v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology mismatch
problem is to construct a minimal spanning tree,  where node degree is
restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
\end{definition}

Early incarnations of the overlay protocols, however, did not make use of the
optimal mapping with the physical network topology. Gnutella protocol for
example was considered far from scalable \cite{ritter_gnucantscale_2001}. Since
every peer can chοse its logical neighbours without knowledge of the underlying
network had as a result a mismatch between physical and application-level
topologies. Additionally, queries may be flooded to multiple paths merging in
one peer while in such case one of the paths would have been enough.
Nevertheless, peers may forward the same message to each other before they
receive the query messages from each other. \cite{matei_mapgnutella_2002} showed
that, even given that 95\% of any two nodes are less than 7 hops away from each
other, a flooding-based algorithm, like the one used by Gnutella, can generate
330TB/month in a 50.000 node network! Therefore, P2P traffic can have a serious
effect on the overall well-being of the Internet as a whole.

Similar problems are observed in decentralized structured schemes also.
Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
balancing, scalability and robustness. Unfortunately, this randomness has a
negative impact on the \emph{routing locality} of the network. This means that
even though the target node can be reached with logarithmic overlay hops, the
distance traveled in the physical underlying network, during the overlay routing
process, can be far from optimal.  The ratio of the physical distance, a query
travels through the overlay to an object, and the minimal distance to that
object (i.e. through IP) is known in the literature as \emph{stretch} or
\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Motivation and our Goal for this Survey}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A topology unaware overlay network is able to control the sequence of peers a
message traverses before reaching its destination, but it is completely unaware
of how the actual packets are switched at the underlying infrastructure along
the overlay path. For example, a single logical point-to-point link on the
overlay, most of the time, corresponds to multiple physical links in the
underlying layer. Additionally, a link in the underlying network often serves
the mapping of several overlay paths causing increase of the traffic on the
physical link, which is also called the link's \emph{stress}\cite{chu_esm_2002}.
Furthermore, the network's \emph{churn}, that is the stochastic behaviour of
peers randomly joining and leaving the P2P network, all cause the overlay to
physical mapping to create such unnecessary, redundant maintenance traffic that
impacts the efficiency of the network in terms of average response time.

The topology mismatch problem is exacerbated by the high amounts of traffic
generated by the highly popular P2P applications. This creats the killer
combination for the network infrastructure, ultimately effecting all Internet
users. Solving it, on the other hand, is no easy task. Topology mismatch is
known to be an NP-Hard problem \cite{NPBOOK,chawathe_scattercast_2000}.
Moreover, the Internet itself is an interesting environment where end-to-end
latencies demonstrate what is called \emph{triangle inequality violations
(TIVs)}, which further complicate the problem. These delay space TIV is a
consequence of the Internet's structure and routing policies and thus will
remain a property of the Internet for the foreseeable future
\cite{zheng_irprtt_2005}. TIVs affect network coordinate
\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
and makes the construction of location/delay aware overlays difficult. An
extensive array of heuristic approaches have been proposed in an effort to find
an algorithm that performs reasonably and consistently well. The purpose of this
survey is to gather the extensive knowledge that has been produced recently in
this research field and organise the available information ultimately helping
the reader sift through the myriad of these efforts, understand the different
approaches and identify though a taxonomy the advantages and disadvantages of
each of them.

>>>>

\section{Background}
\label{section:background}

%The abstraction of the overlay network has been proposed as a way to implement
%efficient, fully distributed, application-layer services such as  
%routing messages to destinations that are not known in advance or building
%one-to-many addressing schemes 
% TODO find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
%\cite{} that provide quality-of-service (QoS) guarantees. \emph{IP
%multicast}, as well as other proposals such as \emph{IntServ} and
%\emph{DiffServ} \cite{cisco_diffserv_2005}, have still not seen wide deployment
%largely due to the fact they require support from the IP
%infrastructure. On the other hand, an overlay network can be deployed on
%end-systems running the overlay protocol software, without the cooperation of ISPs.
%Academic research includes
%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.

In this section, we provide some background information on the types
of P2P overlay architectures that have been implemented and deployed.  We also
describe and motivate more formally the topology mismatch problem tackled
by many P2P researchers over the past several years.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peer-to-Peer (P2P) Overlay Architectures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Circa 2000, the first P2P file-sharing systems introduced the notion 
of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends
the notions of server and client to denote the twofold role of the P2P network
participants. Such serverless systems have proven to be able to achieve outstanding
aggregate resource capacities as more and more participants join the system
without requiring additional expenditure for infrastructure.  While
certain undesirable features such as 
free-riding~\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005},
the distribution of illegal content and/or other \emph{socio-technical}
issues \cite{hughes_socp2p_2008}, have arisen, P2P file-sharing systems 
have continued
to gain popularity because they have successfully supported
popular
applications amongst vast numbers of participating
users.

For over a decade, excitement over the immense potential of P2P systems
has generated a flurry of research and development, resulting in a wide 
range of popular protocols, networks, and applications. As P2P systems evolved
three main architectures have been emerged, namely
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{centralized},
  \item \emph{decentralized unstructured}, and
  \item \emph{decentralized structured}.
\end{inparaenum}

Designers of \emph{centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised})
were the first to recognize that requests for resources (e.g. CPU 
cycles~\cite{seti} or popular content
% TODO find citation for Napster
\cite{}) need not be sent to a dedicated server but instead could be handled by
the many hosts that already posess the resources. The downside of
this approach is it requires a centralized searching facility which becomes
a single point of failure, a scalability bottleneck, and a target of malicious 
attacks (e.g., \emph{Denial-of-Service (DoS) attacks}).

The most successful incarnation of the centralized approach is
the file-sharing application \emph{Napster}
% TODO find citation for Napster again
\cite{}.
Napster maintained a \emph{central index server} based on file lists provided by
participating peers. The central index server was queried by users and it
returned pointers to the actual content. Thus, by centralizing search while
distributing downloads, Napster achieved a highly functional design that 
at its height, attracted 26 million users sharing 80 million 
songs~\cite{jmm_naptopusage_2001}.  
Napster's centralized nature was ultimately what brought it down.
The Recording Industry Association of
America (RIAA) 
filed a lawsuit for copyright infringement of the intellectual
property of artists whose music was distributed by the Napster application.
RIAA was able to point to the centralized indexing component of
Napster as the entity responsible for illegal song-trading.

The centralized approach was quickly replaced by architectures that
distribute both search and resource provision capabilities (see
Figures~\ref{figure:p2p-archs:unstructured-full} and
\ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized}
architectures, resource placement within the overlay topology is 
random~\cite{yang_improvep2psearch_2002}. For this
reason they are referred to as \emph{unstructured}. The most important
properties of such systems are that they support inherent heterogeneity of
peers, are highly resilient to peer failures, and incur low maintenance overhead
at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In
the literature, they are also known as \emph{broadcast-based} systems, because
they use message flooding among peers to propagate
search queries. A well-known example of the fully dentralized unstructured 
approach is Gnutella, in its first realisation (v.0.4) \cite{gnutellav04}.

A key disadvantage of the unstructured approach
is that message flooding burdens the network since queries travel within the
network randomly, visiting nodes that do not have the resource of interest
and thus wasting bandwidth.  To save on bandwidth, unstructured networks
typically limit how long a query travels within the overlay (via a
time-to-live (TTL) parameter).   When a query is received by a node, it decreases
the TTL by one unit.  If the TTL has not reached zero, the node
forwards the query to its neighbors.  Otherwise, it drops the query.  While
this approach prevents queries from visiting the entire network, it does
not guarantee that the resource of interest, (e.g., a rare recording 
of a musical group from
the 1930s) will be found.

To improve search scope and scalability and to lighten network load, 
unstructured systems evolved into ``hierarchical'' systems that dynamically
assign indexing functions to special peers,
called \emph{ultra-peers}  or \emph{super-peers}.
Peers with reliable network connections
and/or high compute power became ultra/super-peers facilitating less 
powerful peers (called \emph{leaf-nodes}) to find resources of interest.
Well-known examples of these hierarchical unstructured systems include
Gnutella v.0.6~\cite{gnutella}, Limewire[XXX], and KaZaA[XXX], and Skype[XXX]. 
%% TODO: add references for the above line.
While the hierarchical approach relieves network pressure as search
queries are flooded over a much smaller subset of peers (the ultra/super-peers),
the disadvantages of reduced search scope and inability to locate
rare resources efficiently remain.

To fully address the disadvantages of decentralized unstructured approaches,
\emph{decentralized structured} schemes have been proposed (see
Figure~\ref{figure:p2p-archs:structured}).  The key goal
of these systems is to provide a self-organising
infrastructure for large-scale P2P applications
\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes
through a deterministic mechanism. 
The DHT  
provides a guaranteed bound on the number of overlay routing
hops that have to be taken to locate a resource, even in the
case where only a single copy of the resource exists in the system.  This means
$O \left ( log n \right )$ hops, compared to unstructured networks that require
$O \left ( n \right )$ to reliably locate a specific object.
Unfortunately this efficiency does not come without cost. Structured overlays
can only support exact-match queries (i.e., queries that identify resources
by name) as opposed to content-based
retrieval provided by unstructured overlays.  This means 
that unstructured overlays can support more 
versatile resource location queries such as keyword and partial matching
searches.  Several popular file-sharing peer-to-peer applications have
been based on DHTs including Overnet[], Kademlia[], the Kad Network[], 
some flavors of BitTorrent, and many
more.  Moreover, DHTs have sparked a flurry of research activity and 
development for a large range of applications: from distributed search 
engines to event notification infrastructures to cloud-based platforms.
We cite several examples in the Introduction.

@@

@@

If these peers participate in an overlay network according to one of the setups
of Figure~\ref{figure:overlay-confs}, then users will experience different
performances. In the overlay depicted in Figure~\ref{figure:overlay-1}, the
message will traverse the following sequence of links in the physical layer
(marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2
\rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5
\rightarrow 5 \rightarrow 4$. In the alternative overlay of Figure
~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$. The
total cost of the first sequence is $45 ms$, while the second yields a mere $15
ms$. Therefore, we can conclude that the second overlay is more congruent with
the underlying physical network than the first one and, thus, is more efficient.

@@

Ideally, an overlay network should be constructed to achieve an optimal 
mapping among the
overlay paths and the underlying IP network links to avoid inefficient states.
The problem of constructing an optimal overlay is referred to as the
\emph{topology mismatch problem}, and is formally defined as follows:
\begin{definition}
Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes,
$\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
$v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology mismatch
problem is to construct a minimal spanning tree,  where node degree is
restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
\end{definition}

Early incarnations of the overlay protocols did not make use of the
optimal mapping with the underlying network topology. Gnutella, for
example, is considered far from scalable \cite{ritter_gnucantscale_2001} because
every peer chooses its neighbors in the overlay without any knowledge
of the underlying network, resulting in a mismatch between the two topologies.
Queries may be flooded over multiple paths merging in
one peer while one of the paths would have been enough.
Moreover, peers may forward the same message to each other before they
receive the query messages from each other. \cite{matei_mapgnutella_2002} showed
that, even when 95\% of any two nodes are less than 7 hops away from each
other, a flooding-based algorithm, like the one used by Gnutella, can generate
330TB/month in a 50.000 node network. A topology mismatch can thus have
a serious effect on Internet router performance.

Similar problems are observed in decentralized structured schemes as well.
Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
balancing, scalability, and robustness for the overlay. Unfortunately, this randomness has a
negative impact on the \emph{routing locality} of the network. This means that
even though the target node can be reached with logarithmic overlay hops, the
distance traveled in the underlying network, during the overlay routing
process, can be far from optimal.  The ratio of the actual IP network distance a query
travels to reach an object of interest (via overlay routing) to
the minimal IP network distance to that
object (i.e. through IP) is known as \emph{stretch} or
\emph{Relative Delay Penalty (RDP)} \cite{chu_esm_2000}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Motivation and our Goal for this Survey}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A topology unaware overlay network is able to control the sequence of peers a
message traverses before reaching its destination, but it is completely unaware
of how the actual packets are switched at the underlying infrastructure along
the overlay path. For example, a single logical point-to-point link on the
overlay typically corresponds to multiple physical links in the
underlying layer. Additionally, a link in the underlying network often lies
on several overlay paths causing increase of the traffic on the
link, known as the link's \emph{stress}~\cite{chu_esm_2002}.
XXX MR:  The following sentence is unclear: XXX
Furthermore, the network's \emph{churn}, that is the stochastic behaviour of
peers randomly joining and leaving the P2P network, all cause the overlay to
physical mapping to create such unnecessary, redundant maintenance traffic that
impacts the efficiency of the network in terms of average response time.

The topology mismatch problem is exacerbated by the extremely large
amounts of traffic
generated by highly popular P2P applications. As noted in the Introduction,
studies have found that more than 60\% of the traffic on the Internet is
P2P.  Thus, solving the topology mismatch problem is a fundamental challenge
in contributing to the overall
efficiency of the Internet.  Unfortunately, topology mismatch is
known to be an NP-Hard problem \cite{NPBOOK,chawathe_scattercast_2000}.
Adding to the complexity of the problem is the fact that end-to-end latencies 
demonstrate \emph{triangle inequality violations
(TIVs)} which are a 
consequence of the Internet's structure and routing policies and thus will
remain a property of the Internet for the foreseeable future
\cite{zheng_irprtt_2005}. TIVs affect both network coordinate
\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
systems 
and makes the construction of IP-topology-aware overlays difficult.

Over the past several years, researchers have proposed and studied 
an extensive array of heuristic solutions to the topology 
mismatch problem.
The purpose of this
survey is to gather and organize the extensive knowledge that has been produced. 
We hope to assist the reader
to understand the different approaches that have been proposed to tackle
the topology mismatch problem and to identify, through our taxonomy, the
advantages and disadvantages of each of them.
